Data Analysis

### 2025-07-06 docs/code/07_data_analysis.Rmd
Results comments addressed (initial stage).. continue results and discussion

---
title: "Green Nest Material Meta-Analysis"
subtitle: "The adaptive value of green nest material across birds: A systematic review and meta-analysis"
author: "Shreya Dimri"
date: last-modified

format:
  html:
   toc: true 
   toc-title: "Table of Contents" 
   toc-depth: 2 
   toc-expand: 3 
   toc-location: left 
   mainfont: Verdana 
   monofont: Jetbrains Mono 
   fontsize: 12pt 
   code-fold: show
   df-print: paged 
   highlight-style: github 
   pdf: default
editor: visual
---

## Loading Data for analysis

```{r setup, warning=FALSE, message=FALSE, echo=F}

@@

# For plotting trees, and although not strictly necessary, we may need the `ggtree` R package, which you should be able to install by running:
# if (!require("BiocManager", quietly = TRUE))
# install.packages("BiocManager")
# BiocManager::install("ggtree")

#devtools::install_github("daniel1noble/orchaRd", ref = "main", force = TRUE)
pacman::p_load(here, tidyverse, metafor,ggpubr,orchaRd,ggpubfigs) #Packages needed
rm(list=ls()) ## cleaning up
set.seed(430) #set seed for reproducibility (for random number generation)
# Loading cleaned dataset
dataset_analysis<- read_csv(here::here("data/03_data_cleaning/dataset_analysis.csv"))

@@

# function to estimate typical sampling error variance obtained from 
# https://github.com/Yefeng0920/heterogeneity_ecoevo/tree/main/function
sigma2_v <- function(mod){
  sigma2_v <- sum(1 / mod$vi) * (mod$k - 1) /
    (sum(1 / mod$vi)^2 - sum((1 / mod$vi)^2))
  return(sigma2_v)
}
```

@@

```{r}
# install.packages("pacman")
# load packages
pacman::p_load(rotl,
               ape,
               ggtree)
# Doesn't matter if I take dataset_lnRR or dataset_SMDH
bird_species <- unique(dataset_analysis$bird_species)

# extracting taxonomic information
taxa <- rotl::tnrs_match_names(names = bird_species)
taxa
# check approximate matches
taxa[taxa$approximate_match==TRUE & !(is.na(taxa$approximate_match)),]
# phylogenetic information for our species from the Open Tree of Life
tree <- rotl::tol_induced_subtree(ott_ids = taxa[["ott_id"]], label_format = "name")
ape::is.binary(tree)
# here are the species included in the tree
sort(tree$tip.label)

@@

sankey_data <- dataset_analysis %>%
  select(Observation_ID, repeated_trait_ID_coded, paper_ID, experiment_ID_coded, group_ID_coded)%>%
  mutate(across(c(Observation_ID, repeated_trait_ID_coded, paper_ID, experiment_ID_coded, group_ID_coded), as.character)) %>%
  make_long(Observation_ID, repeated_trait_ID_coded, paper_ID, experiment_ID_coded, group_ID_coded)

sankey_data <- dataset_analysis %>%
  select(paper_ID, bird_species, population_ID)%>%
  mutate(across(c(paper_ID, bird_species, population_ID), as.character)) %>%
  make_long(paper_ID, bird_species, population_ID)

@@

@@

## For lnRR
# selecting dataset which are not flagged
dataset_lnRR<-dataset_analysis%>%
   filter(lnRR_sign != "NA" | lnRR_variance != "NA")%>%
  filter(proxy_decision == "include")

@@

@@

Exploring heterogeneity in intercept only model

This is the function from Yefeng's Github for interpretting heterogeneity. I have to download the data used for this function to be able to use it. For now I am using the values from the table directly so I should just delete it later before finishing the code.

@@

## lnRR-heterogeneity

@@

# To calculate heterogeneity, these are the functions from orchaRd:
# # I2, CV and M
sigma2_main_model_vector <- sum(intercept_lnRR$sigma2)
# 0.11, 0.27, and 0.57 for lnRR
I2_main_model_vector <- orchaRd::i2_ml(intercept_lnRR)
# 0.88, 0.95, and 0.99 for lnRR
CV_main_model_vector <- orchaRd::cvh1_ml(intercept_lnRR)
# CVH2 (CVH1) - 1.36 (1.16), 3.76 (1.94), and 12.1 (3.48) for lnRR
M_main_model_vector <- orchaRd::m1_ml(intercept_lnRR)
# M2(M1) - 0.58 (0.54), 0.79 (0.66), and 0.78 for lnRR;

@@

their sampling variances will be fitted as variance--covariance matrices assuming a 0.5 correlation (ρs) between sampling variances from the same paper_ID (Noble et al. 2017).

@@

# saving the model
# saveRDS(intercept_SMDH, file = here::here("model/intercept_SMDH.rds"))
```

@@

# To calculate heterogeneity, these are the functions from orchaRd:
# # I2, CV and M
sigma2_main_model_vector <- sum(intercept_SMDH$sigma2)
# 0.54, 1.25, and 3.03 for SMD;
I2_main_model_vector <- orchaRd::i2_ml(intercept_SMDH)
# 0.78, 0.89, and 0.96 for SMD
# 0.88, 0.95, and 0.99 for lnRR
CV_main_model_vector <- orchaRd::cvh1_ml(intercept_SMDH)
# CVH2 (CVH1) - 1.36 (1.16), 3.76 (1.94), and 12.1 (3.48) for lnRR
M_main_model_vector <- orchaRd::m1_ml(intercept_SMDH)
# M2 (M1) - 0.58 (0.54), 0.79 (0.66), and 0.78 for lnRR

@@

saveRDS(Bivariate_model, file = here::here("sensitivity_analysis/bivariate_model.rds"))

@@

### Sensitivity Analysis with published data only

@@

@@

# saving the model
saveRDS(SA_lnRR_published, file = here::here("sensitivity_analysis/SA_lnRR_published.rds"))

@@

@@

# saving the model
saveRDS(SA_SMDH_published, file = here::here("sensitivity_analysis/SA_SMDH_published.rds"))

@@

### Sensitivity Analysis with flagged proxies

@@

# saving the model
saveRDS(SA_lnRR_flagged, file = here::here("sensitivity_analysis/SA_lnRR_flagged.rds"))

@@

## Now for SMDH

@@

# saving the model
saveRDS(SA_SMDH_flagged, file = here::here("sensitivity_analysis/SA_SMDH_flagged.rds"))

@@

### Sensitivity Analysis without 0 ES

@@

# saving the model
saveRDS(SA_lnRR_no_missingES, file = here::here("sensitivity_analysis/SA_lnRR_no_missingES.rds"))

@@

## Now we do this for SMDH

@@

# saving the model
saveRDS(SA_SMDH_no_missingES, file = here::here("sensitivity_analysis/SA_SMDH_no_missingES.rds"))

@@

### Sensitivity Analysis with flagged proxies and no missing ES

@@

# saving the model
saveRDS(SA_lnRR_flagged_no_missingES, file = here::here("sensitivity_analysis/SA_lnRR_flagged_no_missingES.rds"))

@@

## with SMDH

@@

# saving the model
saveRDS(SA_SMDH_flagged_no_missingES, file = here::here("sensitivity_analysis/SA_SMDH_flagged_no_missingES.rds"))

@@

### Sensitivity analysis only using effect sizes calculated from means, SDs, and ns

@@

```{r}
# Preparing new variance-covariance matrix

@@

# saving the model
saveRDS(SA_SMDH_no_inferential, file = here::here("sensitivity_analysis/SA_SMDH_no_inferential.rds"))

@@

### Sensitivity Analysis excluding the parasite and pathogen load fitness proxies

## For the lnRR effect size...
dataset_lnRR_no_PCH<-dataset_lnRR%>%filter(Hypothesis!="PCH")

## The dataset is reduced to almost half
nrow(dataset_lnRR)-nrow(dataset_lnRR_no_PCH)

VCV_lnRR_no_PCH <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR_no_PCH,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

SA_lnRR_no_PCH<- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_no_PCH, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_lnRR_no_PCH)

## For the SMDH effect size..

dataset_SMDH_no_PCH<-dataset_SMDH%>%filter(Hypothesis!="PCH")

## The dataset is reduced to almost half
nrow(dataset_SMDH)-nrow(dataset_SMDH_no_PCH)

VCV_SMDH_no_PCH <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_no_PCH,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

SA_SMDH_no_PCH<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_no_PCH, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_SMDH_no_PCH)
```

### Sensitivity Analysis for excluding non-aromatic and blank (the two control group comparison)

SA_lnRR_no_PCH<- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_no_PCH, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_lnRR_no_PCH)
```

### Different values of sampling variance correlation

@@

```{r}
lnRR_hypothesis <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR,# specify lnRR's sampling variance;
                    mods = ~ - 1 + Hypothesis,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR)

saveRDS(lnRR_hypothesis, file = here::here("model/lnRR_hypothesis.rds"))

@@

saveRDS(SMDH_hypothesis, file = here::here("model/SMDH_hypothesis.rds"))
```

# Exploratory Analysis

For our exploratory analyses (see section Exploratory Analyses), we will run unimoderator multilevel meta-regressions using the following moderators:

#### Type of parasites

@@

saveRDS( lnRR_parasite, file = here::here("model/lnRR_parasite.rds"))

@@

saveRDS(SMDH_parasite, file = here::here("model/SMDH_parasite.rds"))
```

#### Time of addition of green nest material

saveRDS(lnRR_time_gnm_addition, file = here::here("model/lnRR_time_gnm_addition.rds"))

saveRDS(SMDH_time_gnm_addition, file = here::here("model/SMDH_time_gnm_addition.rds"))
```

#### Type of experimental design

```{r}
# This analysis only includes dataset that has a proxy relating to comaparision type
dataset_lnRR_design <- dataset_lnRR %>%
  mutate(comparision_type= case_when(is.na(comparision_type) ~ 4, TRUE ~ comparision_type))

VCV_lnRR_design <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR_design,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )
lnRR_design <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_design ,# specify lnRR's sampling variance;
                    mods = ~ - 1 + as.factor(comparision_type),
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR_design )

saveRDS(lnRR_design, file = here::here("model/lnRR_design.rds"))

# This analysis only includes dataset that has a proxy relating to comaparision type
dataset_SMDH_design <- dataset_SMDH %>%
  mutate(comparision_type= case_when(is.na(comparision_type) ~ 4, TRUE ~ comparision_type))

>>>>

---
title: "Data Analysis"
author: "Shreya D."
date: "`r Sys.Date()`"
output: html_document
---

# Loading Data for analysis

```{r load_packages-data, warning=FALSE, message=FALSE}

@@

# For plotting trees, and although not strictly necessary, we may need the `ggtree` R package, which you should be able to install by running:
 if (!require("BiocManager", quietly = TRUE))
 install.packages("BiocManager")
 BiocManager::install("ggtree")

if (!require("orchaRd", quietly = TRUE))
devtools::install_github("daniel1noble/orchaRd", ref = "main", force = TRUE)

pacman::p_load(here,tidyverse,metafor,ggpubr,orchaRd,ggpubfigs)

#Packages needed

rm(list=ls()) ## cleaning up

set.seed(430) #set seed for reproducibility (for random number generation)

@@

# function to estimate typical sampling error variance obtained from 
# https://github.com/Yefeng0920/heterogeneity_ecoevo/tree/main/function
sigma2_v <- function(mod){
  sigma2_v <- sum(1 / mod$vi) * (mod$k - 1) /
    (sum(1 / mod$vi)^2 - sum((1 / mod$vi)^2))
  return(sigma2_v)
}

# ------------------------------------------------------------
# Loading cleaned dataset
# ------------------------------------------------------------

dataset_analysis<-read_csv(here::here("data/04_data_analysis/dataset_analysis.csv"))

dataset_lnRR<-read_csv(here::here("data/04_data_analysis/dataset_lnRR.csv"))

dataset_SMDH<-read_csv(here::here("data/04_data_analysis/dataset_SMDH.csv"))

@@

```{r}
# install.packages("pacman")
# load packages

pacman::p_load(rotl,
               ape,
               ggtree)

# We have 7 bird species in the dataset
bird_species <- unique(dataset_analysis$bird_species)

# extracting taxonomic information
taxa <- rotl::tnrs_match_names(names = bird_species)
taxa

# check approximate matches
taxa[taxa$approximate_match==TRUE & !(is.na(taxa$approximate_match)),]

# phylogenetic information for our species from the Open Tree of Life

tree <- rotl::tol_induced_subtree(ott_ids = taxa[["ott_id"]], label_format = "name")
ape::is.binary(tree)

# here are the species included in the tree
sort(tree$tip.label)

@@

sankey_data <- dataset_analysis %>%
  select(Observation_ID, repeated_trait_ID_coded, paper_ID, experiment_ID_coded, group_ID_coded)%>%
  mutate(across(c(Observation_ID, repeated_trait_ID_coded, paper_ID, experiment_ID_coded, group_ID_coded), as.factor)) %>%
  make_long(Observation_ID, repeated_trait_ID_coded, paper_ID, experiment_ID_coded, group_ID_coded)

sankey_data <- dataset_analysis %>%
  select(paper_ID, bird_species, population_ID)%>%
  mutate(across(c(paper_ID, bird_species, population_ID), as.factor)) %>%
  make_long(paper_ID, bird_species, population_ID)

@@

```{r}
# Let's just quickly ensure all the important variables are 
# in the correct data type

dataset_analysis<- dataset_analysis%>%mutate(across(c(paper_ID, Observation_ID, repeated_trait_ID_coded, paper_ID, experiment_ID_coded, group_ID_coded, comparision_type,Hypothesis,trait_type,parasite_type,time_of_gnm_addition, blinding, random_assignment, missing_data, population_ID,bird_species), as.factor))

dataset_lnRR<- dataset_lnRR%>%mutate(across(c(paper_ID, Observation_ID, repeated_trait_ID_coded, paper_ID, experiment_ID_coded, group_ID_coded, comparision_type,Hypothesis,trait_type,parasite_type,time_of_gnm_addition, blinding, random_assignment, missing_data, population_ID, bird_species), as.factor))

dataset_SMDH<- dataset_SMDH%>%mutate(across(c(paper_ID, Observation_ID, repeated_trait_ID_coded, paper_ID, experiment_ID_coded, group_ID_coded, comparision_type,Hypothesis,trait_type,parasite_type,time_of_gnm_addition, blinding, random_assignment, missing_data, population_ID,bird_species), as.factor))

@@

## For lnRR (or simply load dataset as provided)
# dataset_lnRR<-dataset_analysis%>%
# filter(lnRR_sign != "NA" | lnRR_variance != "NA")

# selecting dataset which are not flagged
dataset_lnRR<-dataset_lnRR%>%
  filter(proxy_decision == "include")

@@

# selecting only dataset that passes geary's test
lnRR_geary_passed<-dataset_lnRR%>%filter(geary_test=="pass")

@@

# saving the dataset for this model
write.csv(dataset_lnRR,here::here("data/04_data_analysis/intercept_lnRR.csv"))
```

#### Exploring heterogeneity in intercept only model

This is the function from Yefeng's Github for interpreting heterogeneity. I have to download the data used for this function to be able to use it.

I am using the values from the table directly..

@@

@@

# To calculate heterogeneity, these are the functions from orchaRd:
# # I2, CV and M
sigma2_main_model_vector <- sum(intercept_lnRR$sigma2)

# 0.11, 0.27, and 0.57 for lnRR

I2_main_model_vector <- orchaRd::i2_ml(intercept_lnRR)

# 0.88, 0.95, and 0.99 for lnRR

CV2_main_model_vector <- orchaRd::cvh2_ml(intercept_lnRR)
CV1_main_model_vector <- orchaRd::cvh1_ml(intercept_lnRR)

# CVH2 (CVH1) - 1.36 (1.16), 3.76 (1.94), and 12.1 (3.48) for lnRR

M2_main_model_vector <- orchaRd::m2_ml(intercept_lnRR)
M1_main_model_vector <- orchaRd::m1_ml(intercept_lnRR)

# M2(M1) - 0.58 (0.54), 0.79 (0.66), and 0.78 for lnRR;

@@

The sampling variances will be fitted as variance--covariance matrices assuming a 0.5 correlation (ρs) between sampling variances from the same paper_ID (Noble et al. 2017).

@@

# saving the model
saveRDS(intercept_SMDH, file = here::here("model/intercept_SMDH.rds"))

# saving the dataset for this model
write.csv(dataset_SMDH,here::here("data/04_data_analysis/intercept_SMDH.csv"))
```

@@

# To calculate heterogeneity, these are the functions from orchaRd:
# # I2, CV and M
sigma2_main_model_vector <- sum(intercept_SMDH$sigma2)
# 0.54, 1.25, and 3.03 for SMD;
I2_main_model_vector <- orchaRd::i2_ml(intercept_SMDH)
# 0.78, 0.89, and 0.96 for SMD
# 0.88, 0.95, and 0.99 for lnRR
CV2_main_model_vector <- orchaRd::cvh2_ml(intercept_SMDH)
CV1_main_model_vector <- orchaRd::cvh1_ml(intercept_SMDH)
# CVH2 (CVH1) - 1.36 (1.16), 3.76 (1.94), and 12.1 (3.48) for lnRR
M2_main_model_vector <- orchaRd::m2_ml(intercept_SMDH)
M1_main_model_vector <- orchaRd::m1_ml(intercept_SMDH)
# M2 (M1) - 0.58 (0.54), 0.79 (0.66), and 0.78 for lnRR

@@

# Saving the model
saveRDS(Bivariate_model, file = here::here("model/sensitivity_analysis/bivariate_model.rds"))
# saving the dataset for this model
write.csv(dataset_bivariate,here::here("data/05_data_sensitivity_analysis/dataset_bivariate.csv"))

@@

### With published data only

@@

#----------------------------------------------------------------------------
# for lnRR as Effect Size
#----------------------------------------------------------------------------

@@

# Saving the model
saveRDS(SA_lnRR_published, file = here::here("model/sensitivity_analysis/SA_lnRR_published.rds"))
# saving the dataset for this model
write.csv(dataset_lnRR_published,here::here("data/05_data_sensitivity_analysis/dataset_lnRR_published.csv"))

@@

#----------------------------------------------------------------------------
# for SMDH as Effect Size
#----------------------------------------------------------------------------

@@

# Saving the model
saveRDS(SA_SMDH_published, file = here::here("model/sensitivity_analysis/SA_SMDH_published.rds"))
# saving the dataset for this model
write.csv(dataset_SMDH_published,here::here("data/05_data_sensitivity_analysis/dataset_SMDH_published.csv"))

@@

### With flagged proxies

Some traits for which the relationship with fitness is unclear or highly debated in the literature

@@

# Saving the model
saveRDS(SA_lnRR_flagged, file = here::here("model/sensitivity_analysis/SA_lnRR_flagged.rds"))
# saving the dataset for this model
write.csv(lnRR_con_flagged,here::here("data/05_data_sensitivity_analysis/dataset_lnRR_con_flagged.csv"))

@@

#----------------------------------------------------------------------------
# for SMDH as Effect Size
#----------------------------------------------------------------------------

@@

# Saving the model
saveRDS(SA_SMDH_flagged, file = here::here("model/sensitivity_analysis/SA_SMDH_flagged.rds"))
# saving the dataset for this model
write.csv(SMDH_con_flagged,here::here("data/05_data_sensitivity_analysis/dataset_SMDH_con_flagged.csv"))

@@

### Without 0 ES

@@

# Saving the model
saveRDS(SA_lnRR_no_missingES, file = here::here("model/sensitivity_analysis/SA_lnRR_no_missingES.rds"))
# saving the dataset for this model
write.csv(lnRR_sin_missingES,here::here("data/05_data_sensitivity_analysis/dataset_lnRR_sin_missingES.csv"))

@@

#----------------------------------------------------------------------------
# for SMDH as Effect Size
#----------------------------------------------------------------------------

@@

# Saving the model
saveRDS(SA_SMDH_no_missingES, file = here::here("model/sensitivity_analysis/SA_SMDH_no_missingES.rds"))
# saving the dataset for this model
write.csv(SMDH_sin_missingES,here::here("data/05_data_sensitivity_analysis/dataset_SMDH_sin_missingES.csv"))

@@

### With flagged proxies and no missing ES

@@

# Saving the model
saveRDS(SA_lnRR_flagged_no_missingES, file = here::here("model/sensitivity_analysis/SA_lnRR_flagged_no_missingES.rds"))

# saving the dataset for this model
write.csv(lnRR_flagged_no_missingES,here::here("data/05_data_sensitivity_analysis/dataset_lnRR_flagged_no_missingES.csv"))

@@

#----------------------------------------------------------------------------
# for SMDH as Effect Size
#----------------------------------------------------------------------------

@@

# Saving the model
saveRDS(SA_SMDH_flagged_no_missingES, file = here::here("model/sensitivity_analysis/SA_SMDH_flagged_no_missingES.rds"))

# saving the dataset for this model
write.csv(SMDH_flagged_no_missingES,here::here("data/05_data_sensitivity_analysis/dataset_SMDH_flagged_no_missingES.csv"))

@@

### No inferential stats for SMDH: effect sizes calculated from means, SDs, and sample size

@@

#----------------------------------------------------------------------------
# for SMDH as Effect Size
#----------------------------------------------------------------------------

@@

# Saving the model
saveRDS(SA_SMDH_no_inferential, file = here::here("model/sensitivity_analysis/SA_SMDH_no_inferential.rds"))

# saving the dataset for this model
write.csv(SMDH_no_inferential,here::here("data/05_data_sensitivity_analysis/dataset_SMDH_no_inferential.csv"))

@@

### Exluding failed geary's test for lnRR

```{r}
#----------------------------------------------------------------------------
# for  lnRR  as Effect Size
#----------------------------------------------------------------------------

VCV_lnRR_geary_passed <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = lnRR_geary_passed,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

SA_lnRR_geary_passed<- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_geary_passed, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = lnRR_geary_passed)

# Saving the model
saveRDS(SA_lnRR_geary_passed, file = here::here("model/sensitivity_analysis/SA_lnRR_geary_passed.rds"))

# saving the dataset for this model
write.csv(lnRR_geary_passed,here::here("data/05_data_sensitivity_analysis/dataset_lnRR_geary_passed.csv"))

### Excluding the parasite and pathogen load fitness proxies

It excludes all the fitness proxies relating to parasite and pathogen loads/ abundance etc.

```{r}
#----------------------------------------------------------------------------
# for  lnRR  as Effect Size
#----------------------------------------------------------------------------

dataset_lnRR_no_para<-dataset_lnRR%>%filter(is.na(parasite_type))

## The dataset is reduced to almost half
nrow(dataset_lnRR)-nrow(dataset_lnRR_no_para)

VCV_lnRR_no_para <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR_no_para,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

SA_lnRR_no_para<- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_no_para, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_lnRR_no_para)

# Saving the model
saveRDS(SA_lnRR_no_para, file = here::here("model/sensitivity_analysis/SA_lnRR_no_para.rds"))

# saving the dataset for this model
write.csv(dataset_lnRR_no_para,here::here("data/05_data_sensitivity_analysis/dataset_lnRR_no_para.csv"))

#----------------------------------------------------------------------------
# for  SMDH  as Effect Size
#----------------------------------------------------------------------------

dataset_SMDH_no_para<-dataset_SMDH%>%filter(is.na(parasite_type))

## The dataset is reduced to almost half
nrow(dataset_SMDH)-nrow(dataset_SMDH_no_para)

VCV_SMDH_no_para <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_no_para,
                  rho = 0.5) # assuming that the effect sizes within the same study are correlated with rho = 0.5

SA_SMDH_no_para<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_no_para, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_SMDH_no_para)

# Saving the model
saveRDS(SA_SMDH_no_para, file = here::here("model/sensitivity_analysis/SA_SMDH_no_para.rds"))

# saving the dataset for this model
write.csv(dataset_SMDH_no_para,here::here("data/05_data_sensitivity_analysis/dataset_SMDH_no_para.csv"))

## Different values of sampling variance correlation

@@

```{r hypothesis-model}
lnRR_hypothesis <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR,# specify lnRR's sampling variance;
                    mods = ~ - 1 + Hypothesis,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR)
# Save model
saveRDS(lnRR_hypothesis, file = here::here("model/lnRR_hypothesis.rds"))

# saving the dataset for this model
write.csv(dataset_lnRR,here::here("data/04_data_analysis/hypothesis_lnRR.csv"))

@@

# Save model
saveRDS(SMDH_hypothesis, file = here::here("model/SMDH_hypothesis.rds"))

# saving the dataset for this model
write.csv(dataset_SMDH,here::here("data/04_data_analysis/hypothesis_SMDH.csv"))
```

# Meta-regressions

For understanding the heterogeneity in our intercept-only model, we will run unimoderator multilevel meta-regressions (see section Exploratory Analyses in pre-registration or manuscript) using the following moderators:

## Type of experimental design

```{r}
# This analysis only includes dataset that has a proxy relating to comaparision type
dataset_lnRR_design <- dataset_lnRR %>% filter(!is.na(comparision_type))

VCV_lnRR_design <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR_design,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )
lnRR_design <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_design ,# specify lnRR's sampling variance;
                    mods = ~ - 1 + comparision_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR_design )

# Saving the model
saveRDS(lnRR_design, file = here::here("model/lnRR_design.rds"))

# saving the dataset for this model
write.csv(dataset_lnRR_design,here::here("data/04_data_analysis/design_lnRR.csv"))

# This analysis only includes dataset that has a proxy relating to comaparision type
dataset_SMDH_design <- dataset_SMDH %>% filter(!is.na(comparision_type))

VCV_SMDH_design <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_design,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )
SMDH_design <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_design ,# specify lnRR's sampling variance;
                    mods = ~ - 1 + comparision_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH_design )

# Saving the model
saveRDS(SMDH_design, file = here::here("model/SMDH_design.rds"))

# saving the dataset for this model
write.csv(dataset_SMDH_design,here::here("data/04_data_analysis/design_SMDH.csv"))
```

### Excluding non-aromatic and blank comparison (i.e. the two control group comparison)

SA_lnRR_no_comp3<- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_no_comp3, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_lnRR_no_comp3)

# Saving the model
saveRDS(SA_lnRR_no_comp3, file = here::here("model/sensitivity_analysis/SA_lnRR_no_comp3.rds"))

# saving the dataset for this model
write.csv(dataset_lnRR_no_comp3,here::here("data/05_data_sensitivity_analysis/dataset_lnRR_no_comp3.csv"))

#----------------------------------------------------------------------------
# for  SMDH  as Effect Size
#----------------------------------------------------------------------------

dataset_SMDH_no_comp3<-dataset_SMDH%>%filter(comparision_type!=3)
## Only some rows are removed... 
nrow(dataset_SMDH)-nrow(dataset_SMDH_no_comp3)

VCV_SMDH_no_comp3 <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_no_comp3,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

SA_SMDH_no_comp3<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_no_comp3, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_SMDH_no_comp3)

# Saving the model
saveRDS(SA_SMDH_no_comp3, file = here::here("model/sensitivity_analysis/SA_SMDH_no_comp3.rds"))

# saving the dataset for this model
write.csv(dataset_SMDH_no_comp3,here::here("data/05_data_sensitivity_analysis/dataset_SMDH_no_comp3.csv"))
```

## Time of addition of green nest material

## Saving the model
saveRDS(lnRR_time_gnm_addition, file = here::here("model/lnRR_time_gnm_addition.rds"))

# saving the dataset for this model
write.csv(dataset_lnRR,here::here("data/04_data_analysis/time_gnm_lnRR.csv"))

## Saving the model
saveRDS(SMDH_time_gnm_addition, file = here::here("model/SMDH_time_gnm_addition.rds"))

# saving the dataset for this model
write.csv(dataset_SMDH,here::here("data/04_data_analysis/time_gnm_SMDH.csv"))
```

## Type of parasites

@@

# Saving the model
saveRDS(lnRR_parasite, file = here::here("model/lnRR_parasite.rds"))

# saving the dataset for this model
write.csv(dataset_lnRR_parasite,here::here("data/04_data_analysis/parasite_lnRR.csv"))

@@

# Save the model
saveRDS(SMDH_parasite, file = here::here("model/SMDH_parasite.rds"))

# saving the dataset for this model
write.csv(dataset_SMDH_parasite,here::here("data/04_data_analysis/parasite_SMDH.csv"))
```
